Charged pion emission from neutron-rich heavy ion collisions for studies on the symmetry energy

### 2020-08-07 chapters_altFormat/ch_data_analysis_2.tex
Put in Pawel and Vladimir edits

\begin{equation}
  f(x,y)=\frac1{2\pi\sigma_x\sigma_y\sqrt{1-\rho^2}}\exp\left\{
  \frac{-(x - \mu_{x})^2/\sigma_x^2-(y-\mu_{y})^2/\sigma_y^2+2\rho
  xy/\sigma_x\sigma_y}{2(1-\rho^2)}\right\},
   \label{multiGauss}
\end{equation}
where x=A/Q, y=Z, $\mu$ is the mean values, and $\sigma$ are the Gaussian widths of the two variables. The gates drawn in Figure~\ref{fig:beampidTwo} are summarized in Table~\ref{beamParameters}.

@@

Near the edges of the detection volume, the clusters of tracks significantly deviate from the trend of the fitted track as seen in different view of the reconstructed clusters in Figure~\ref{fig:edge}.  This happens because the last pads on the edges of the pad-plane have no neighboring pads containing charge, and therefore the last pad represents the last known position of the collected charge. In this case the cluster position is biased towards the inside of the TPC. This also occurs for the first and last time bucket in the vertical direction. While the number of affected clusters is small, compared with the total number in the track, the deviation at the end of the track is enough cause issues in the momentum reconstruction. Simple cuts were taken to graphically remove these clusters around the left,right, top, and bottom of the TPC. The hits that were cut out satisfied the following conditions,

@@

 The target position was physically expected to be \SI{-13.2}{\milli\metre} in the TPC coordinate frame. The z-component of all runs in each beam type are plotted around the the target region in Figure~\ref{fig:vertexdist}. The mean position of the vertex from the reconstructed data is around \SI{-14.76}{\milli\metre} which is about \SI{1.6}{\milli\metre} off from the expected target position. The mean position for each system is listed in Table~\ref{tb:vertexresol}. Since the target thickness was less than \SI{1}{\milli\metre} for all targets, we can assume the thickness of the target has a minimal effect on the vertex resolution of the TPC. Therefore the measured width of the vertex distribution can be interpreted directly as the vertex resolution of the TPC. The extracted vertex resolutions of each system are summarized in Table~\ref{tb:vertexresol}, with an average vertex resolution of \SI{1.2}{\centi\metre}. The difference between the measured and actual target location is 10 times smaller than the intrinsic vertex resolution of the detector and is an insignificant difference.

@@

The impact parameter is a theoretical quantity defined as the distance between two centers of colliding nuclei. Though there is no way to directly measure the impact parameter in an experiment, the track multiplicity is indirectly related to the estimated impact parameter of a collision \cite{impactpar}. As discussed in Section \ref{sec:kyoto}, the Kyoto multiplicity array was used to experimentally trigger on central nuclear collision events. This is under the assumption that number of charged particles produced in collision is related to the overlap region of the two nuclei. This is best described by a spectator-participant model of nuclear collisions, where a large fraction of the participating nucleons in the overlap region of the two colliding nuclei fragment into individual and clusters of nucleons. In the case where the impact parameter is zero, all the nucleons in both nuclei participate, were as larger impact parameters -- more peripheral collisions -- less nucleons participate.

@@

where $b_{\mathrm{max}}$ represents the maximum impact parameter detected by the TPC, $b$ is the impact parameter of the vent,  and $dP(N_C/dN_C$ is the normalized multiplicity distribution \cite{reducedimpact}. The reduced impact parameter ranges from $\hat{b}=0$ for the most central collisions and to $\hat{b}=1$ for the most peripheral. A  detailed analysis was performed, determining the maximum cross section $\sigma_{max}$, for each system, in which $b_{\mathrm{max} = \sqrt{\sigma_{max}/\pi}}$. The detailed analysis is given in \cite{jon}. Figure~\ref{fig:impactPar} shows the estimated impact parameter for a given track multiplicity with the estimated error bands for each system.

@@

where $\Delta\phi = (\phi_2 - \phi_1)$  and $\theta_{1,2}$ are the $\theta$ angle cuts, in units of \si{\radian}. The solid angle covered by the $\pi^-$ cuts is \SI{2.077}{\steradian} and for $\pi^+$ cuts \SI{2.042}{\steradian}. Assuming the pion emission is isotropic, we multiply each observed pion by a correction factor, $C_a$, to correct for the full 4$\pi$ angular coverage,

@@

Once all the events have been rotated to align with the z-axis, transforming from the Lab to the CM frame is done by a Lorentz transformation. The 4-momentum vector in the lab frame is defined as $\textbf{P} = (E/c,p_x,p_y,p_z)$. Where the corresponding Lorentz transform into the CM frame along the beam (z-axis) is defined as the Lorentz transformation matrix $A$,

@@

where $\beta$, describes the velocity of the CM system, and $\gamma=\sqrt{1-\beta^2}^{-1}$. The parameter $\beta$ can be determined from the total momentum of the system in the Laboratory frame $P = \sqrt{ T_{P}^2 + 2M_{T}T_{T}}$ and the total energy of the system $E = T_{P} + M_{P} + M_{T}$ -- where $T_{P}$ is the projectile kinetic energy in the Lab frame, $M_{T}$, and $M_{P}$ are the total mass of the target and projectile, where $\beta = -P/E$; the (-) sign denotes the correct direction for transforming from the Lab to the CM frame. The transformation of each  track is defined as $p^{CM} = \textbf{A}p^{Lab}$.

@@

where $P(\pi)$ is the probability that a track is a pion described in Eq.~\ref{eq:pionProbs}, $\epsilon_i(p_{Lab}, \theta_{Lab})$ is the efficiency of particle $i$ given by its momentum $p_{Lab}$ and polar angle $\theta_{Lab}$ as discussed in Section~\ref{sec:efficiency}. The factor $A$ is a scale factor defined as,

@@

\begin{equation}
{\delta \epsilon }^2 = A^2 \sum_i^{N_{\pi}} \left[ \left( \frac{P(\pi)^{\pi} }{\epsilon_i} \right)^2  \left(  \frac{P(\pi)^{\pi} }{\epsilon_i} \frac{\delta \epsilon_i}{\epsilon_i} \right)^2   \right]
\end{equation}

>>>>

\begin{equation}
  f(x,y)=\frac1{2\pi\sigma_x\sigma_y\sqrt{1-\rho^2}}\exp\left\{
  \frac{-(x - \mu_{x})^2/\sigma_x^2-(y-\mu_{y})^2/\sigma_y^2+2\rho
  xy/\sigma_x\sigma_y}{2(1-\rho^2)}\right\},
   \label{multiGauss}
\end{equation}
where x=A/Q, y=Z, $\mu$ are the mean values, and $\sigma$ are the Gaussian widths of the two variables. The gates drawn in Figure~\ref{fig:beampidTwo} are summarized in Table~\ref{beamParameters}.

@@

Near the edges of the detection volume, the clusters of tracks significantly deviate from the trend of the fitted track as seen in different view of the reconstructed clusters in Figure~\ref{fig:edge}.  This happens because the last pads on the edges of the pad-plane have no neighboring pads containing charge, and therefore the last pad represents the last known position of the collected charge. In this case the cluster position is biased towards the inside of the TPC. This also occurs for the first and last time bucket in the vertical direction. While the number of affected clusters is small, compared with the total number in the track, the deviation at the end of the track is enough to cause issues in the momentum reconstruction. Simple cuts were taken to graphically remove these clusters around the left,right, top, and bottom of the TPC. The hits that were cut out satisfied the following conditions,

@@

 The target position was physically expected to be \SI{-13.2}{\milli\metre} in the TPC coordinate frame. The z-component of all runs in each beam type are plotted around the target region in Figure~\ref{fig:vertexdist}. The mean position of the vertex from the reconstructed data is around \SI{-14.76}{\milli\metre} which is about \SI{1.6}{\milli\metre} off from the expected target position. The mean position for each system is listed in Table~\ref{tb:vertexresol}. Since the target thickness was less than \SI{1}{\milli\metre} for all targets, we can assume the thickness of the target has a minimal effect on the vertex resolution of the TPC. Therefore the measured width of the vertex distribution can be interpreted directly as the vertex resolution of the TPC. The extracted vertex resolutions of each system are summarized in Table~\ref{tb:vertexresol}, with an average vertex resolution of \SI{1.2}{\centi\metre}. The difference between the measured and actual target location is 10 times smaller than the intrinsic vertex resolution of the detector and is an insignificant difference.

@@

The impact parameter is a theoretical quantity defined as the distance between two centers of colliding nuclei if they were to follow a straight line path. Though there is no way to directly measure the impact parameter in an experiment, the track multiplicity is indirectly related to the estimated impact parameter of a collision \cite{impactpar}. As discussed in Section \ref{sec:kyoto}, the Kyoto multiplicity array was used to experimentally trigger on central nuclear collision events. This is under the assumption that number of charged particles produced in collision is related to the overlap region of the two nuclei. This is best described by a spectator-participant model of nuclear collisions, where a large fraction of the participating nucleons in the overlap region of the two colliding nuclei fragment into individual and clusters of nucleons. In the case where the impact parameter is zero, all the nucleons in both nuclei participate, were as larger impact parameters -- more peripheral collisions -- less nucleons participate.

@@

where $b_{\mathrm{max}}$ represents the maximum impact parameter detected by the TPC, $b$ is the impact parameter of the event,  and $dP(N_C/dN_C$ is the normalized multiplicity distribution \cite{reducedimpact}. The reduced impact parameter ranges from $\hat{b}=0$ for the most central collisions and to $\hat{b}=1$ for the most peripheral. A  detailed analysis was performed, determining the maximum cross section $\sigma_{max}$, for each system, in which $b_{\mathrm{max}} = \sqrt{\sigma_{max}/\pi}$. The detailed analysis is given in \cite{jon}. Figure~\ref{fig:impactPar} shows the estimated impact parameter for a given track multiplicity with the estimated error bands for each system.

@@

where $\Delta\phi = (\phi_2 - \phi_1)$  and $\theta_{1,2}$ are the $\theta$ angle cuts, in units of \si{\radian}. The solid angle covered by the $\pi^-$ cuts is \SI{2.077}{\steradian} and for $\pi^+$ cuts \SI{2.042}{\steradian}. Assuming the pion emission is isotropic, we weight each observed pion by a correction factor, $C_a$, to correct for the full 4$\pi$ angular coverage,

@@

Once all the events have been rotated to align with the z-axis, transforming from the Lab to the CM frame is done by a Lorentz transformation. The 4-momentum vector in the lab frame is defined as $\textbf{P} = (E/c,p_x,p_y,p_z)$. Where the corresponding Lorentz transform into the CM frame along the beam (z-axis) is defined by the matrix $A$,

@@

where $\beta$, describes the velocity of the CM system, and $\gamma=1/ \sqrt{1-\beta^2}$. The parameter $\beta$ can be determined from the total momentum of the system in the Laboratory frame $P = \sqrt{ T_{P}^2 + 2M_{T}T_{T}}$ and the total energy of the system $E = T_{P} + M_{P} + M_{T}$ -- where $T_{P}$ is the projectile kinetic energy in the Lab frame, $M_{T}$, and $M_{P}$ are the total mass of the target and projectile, where $\beta = -P/E$; the (-) sign denotes the correct direction for transforming from the Lab to the CM frame. The transformation of each  track is defined as $p^{CM} = \textbf{A}p^{Lab}$.

@@

where $P(\pi)$ is the probability that a given track is a pion described in Eq.~\ref{eq:pionProbs}, $\epsilon_i(p_{Lab}, \theta_{Lab})$ is the efficiency of particle $i$ given by its momentum $p_{Lab}$ and polar angle $\theta_{Lab}$ as discussed in Section~\ref{sec:efficiency}. The factor $A$ is a scale factor defined as,

@@

\begin{equation}
{\delta \epsilon }^2 = A^2 \sum_i^{N_{\pi}} \left[ \left( \frac{P(\pi)^{\pi} }{\epsilon_i} \right)^2  \left(  \frac{P(\pi)^{\pi} }{\epsilon_i} + \frac{\delta \epsilon_i}{\epsilon_i} \right)^2   \right]
\end{equation}

### 2020-08-07 chapters_altFormat/ch_experiment.tex
Put in Pawel and Vladimir edits

Figure~\ref{fig:TPC_schematic} shows the schematic of the effective resistances and capacitance of the TPC subsystem. The cathode is connected to the high voltage (HV) supply through a \SI{10}{\mega\ohm} resistor and has an effective capacitance to ground of \SI{4}{\nano\farad}, $C_{VSD}$. The cathode voltage $V_{cath}$ can be calculated as,
\begin{equation}
V_{cath} = \frac{V_{HV}}{ 1 + \frac{10}{ \left( (245 + R_p)^{-1} + 700^{-1} \right)^{-1} } },
\end{equation}
where 
\begin{equation}
R_p = \left( R_{TP1}^{-1} + R_{TP2}^{-1} \right)^{-1},
 \label{eq:Reff}
\end{equation}

@@

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{gg_onoff.pdf}
\caption{Garfield simulations of the on and off configuration of the gating grid.}
\label{fig:gg_onoff}
\end{figure}

Both configurations of the gating grid were measured and simulated. To measure the electron transparency in the on configuration all wires were set to the same voltage $V_{avg}$, and the anode wires were lowered to \SI{500}{\volt}. By lowering the voltage of the anode wires we could measure the large charge of the beam without saturating the electronics. The average charge deposited in the chamber was measured as a function of $V_{avg}$. Changing the average voltage required a different top perimeter resistor value according to Eq.~\ref{eq:TP_resistor}. Several runs were taken ranging from \SIrange{-198}{-40}{\volt}, with and without the magnetic field. Theoretically the most negative value represents 100\% electron transparency and was used as the reference run in which we scaled all measurements to. The measured electron transparency, $T$, was estimated as $T = \langle dE/dx\rangle/{\langle dE/dx\rangle}_{ref}$, where $\langle dE/dx\rangle_{ref}$ represents the average energy loss of the reference run. Figure~\ref{fig:ggAvgTrans} shows the measured transparency as a function of $V_{avg}$, as compared with the corresponding Garfield simulation \cite{garfield++}. The average gating grid voltage used in the experiment was \SI{-171}{\volt} to ensure we were well within the 100\% transparency region.

@@

The Katana trigger box took some time to completely master, it did not take on the major trigger responsiblities at first. In later runs, most of the trigger functions were performed the in Katana trigger box. The master trigger for the DAQ was different for each primary beam as the experiment got progressively better. During the ${}^{124}$Xe primary beam, the KATANA trigger box was an input into the trigger logic, where as in the ${}^{238}$U primary beam, the KATANA trigger box functioned as the trigger logic utilizing the internal trigger electronics. In either case the differences in the trigger logic were very minor and they both behaved practically the same except for minor details on the gating-grid trigger \cite{jon}. Figure~\ref{fig:katanaLogic} summarizes the KATANA trigger box logic.

>>>>

Figure~\ref{fig:TPC_schematic} shows the schematic of the effective resistances of the TPC subsystem. The cathode is connected to the high voltage (HV) supply through a \SI{10}{\mega\ohm} resistor and has an effective capacitance to ground of \SI{4}{\nano\farad}, $C_{VSD}$. The cathode voltage $V_{cath}$ can be calculated as,
\begin{equation}
V_{cath} = \frac{V_{HV}}{ 1 + \frac{10}{ \left( (245 + R_p)^{-1} + 700^{-1} \right)^{-1} } },
\end{equation}
where 
\begin{equation}
R_p = \left( R_{TP1}^{-1} + R_{TP2}^{-1} \right)^{-1},
 \label{eq:Reff}
\end{equation}

@@

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{gg_onoff.pdf}
\caption{Garfield simulations of the on and off configuration of the gating grid with the drift lines of electrons through all of the grids.}
\label{fig:gg_onoff}
\end{figure}

Both configurations of the gating grid were measured and simulated. To measure the electron transparency in the on configuration all wires were set to the same voltage $V_{avg}$, and the anode wires were lowered to \SI{500}{\volt}. By lowering the voltage of the anode wires we could measure the large charge of the beam without saturating the electronics. The average charge deposited in the chamber was measured as a function of $V_{avg}$. Changing the average voltage required a different top perimeter resistor value according to Eq.~\ref{eq:TP_resistor}. Several runs were taken ranging from \SIrange{-198}{-40}{\volt}, with and without the magnetic field. Theoretically the most negative value represents 100\% electron transparency and was used as the reference run in which we scaled all measurements to. The measured electron transparency, $T$, was estimated as $T_i = \langle dE/dx\rangle_i /{\langle dE/dx\rangle}_{ref}$, where $\langle dE/dx\rangle_{ref}$ represents the average energy loss of the reference run, and $\langle dE/dx\rangle_i$ is the average energy loss for the $i^{th}$ run. Figure~\ref{fig:ggAvgTrans} shows the measured transparency as a function of $V_{avg}$, as compared with the corresponding Garfield simulation \cite{garfield++}. The average gating grid voltage used in the experiment was \SI{-171}{\volt} to ensure we were well within the 100\% transparency region.

@@

The Katana trigger box took some time to completely master, it did not take on the major trigger responsiblities at first. In later runs, most of the trigger functions were performed in the Katana trigger box. The master trigger for the DAQ was different for each primary beam. During the ${}^{124}$Xe primary beam, the KATANA trigger box was an input into the trigger logic, whereas in the ${}^{238}$U primary beam, the KATANA trigger box functioned as the trigger logic utilizing the internal trigger electronics. In either case the differences in the trigger logic were very minor and they both behaved practically the same except for minor details on the gating-grid trigger \cite{jon}. Figure~\ref{fig:katanaLogic} summarizes the KATANA trigger box logic.
